Validation of Visual Statistical Inference, Applied to Linear Models

### 2012-06-20 paper_jasa.tex
p-value plot updated
started discussion of subject specific variation

We fit model \ref{mixed} with effect $E=\beta*n/\sqrt(n)$ as the only fixed effect covariate without intercept. We also set $offset = log(0.05/95)$ so that the estimated power has a fixed lower limit = 0.05 (Type-I error) for effect $E=0$. The power curves obtained by fitting model \ref{mixed} to the data for all the experiments applying different screening criteria is shown in figure \ref{fig:power_screening}. This shows how the screening criteria may affect the results. The corresponding power curves of conventional test are also shown. Notice that no mater what criteria we apply the result does not change much for experiment 2. Also, estimated power curve with criteria 5 for experiment 3 is above the conventional test power curve.

@@

\begin{figure*}[hbtp]
   \centering
       \scalebox{0.45}{\includegraphics{effect_power_exp1.pdf}}
       \scalebox{0.45}{\includegraphics{effect_power_exp2.pdf}}
       \caption{Observed power shown by loess smoother and the power of conventional test for different effect size $E$ as per experiment 1 and 2.}
       \label{fig:power_effect1}
\end{figure*}

@@

\begin{figure*}[hbtp]
   \centering
       \scalebox{0.4}{\includegraphics{power_loess_exp3.pdf}}
       \caption{Observed power shown by loess smoother with simultaneous bootstrap confidence band and the power of conventional test for sample size $n= 100$ and $\sigma = 5$ as per experiment 3. Notice that for experiment 3, power does not seem to depend on the parameter $\beta$. Di/Heike, I am not sure how I can relate this to demonstrate the power of visual inference when normal conditions are not met. Does this really test our hypothesis $H_0: \beta=0$ or some other hypothesis about the existence of cluster in the data?}
       \label{fig:power_loess3}
\end{figure*}

@@

>>>>

We fit model \ref{mixed} with effect $E=\beta*n/\sqrt(n)$ as the only fixed effect covariate without intercept. We also set $offset = log(0.05/0.95)$ so that the estimated power has a fixed lower limit = 0.05 (Type-I error) for effect $E=0$. The power curves obtained by fitting model \ref{mixed} to the data for all the experiments applying different screening criteria is shown in figure \ref{fig:power_screening}. This shows how the screening criteria may affect the results. The corresponding power curves of conventional test are also shown. Notice that no mater what criteria we apply the result does not change much for experiment 2. Also, estimated power curve with criteria 5 for experiment 3 is above the conventional test power curve.

@@

%\begin{figure*}[hbtp]
%   \centering
%       \scalebox{0.45}{\includegraphics{effect_power_exp1.pdf}}
%       \scalebox{0.45}{\includegraphics{effect_power_exp2.pdf}}
%       \caption{Observed power shown by loess smoother and the power of conventional test for different effect size $E$ as per experiment 1 and 2.}
%       \label{fig:power_effect1}
%\end{figure*}

@@

\begin{figure*}[hbtp]
   \centering
       \scalebox{0.4}{\includegraphics{power_loess_exp3.pdf}}
       \caption{Observed power shown by loess smoother with simultaneous bootstrap confidence band and the power of conventional test for sample size $n= 100$ and $\sigma = 5$ as per experiment 3. Notice that for experiment 3, power does not seem to depend on the parameter $\beta$. \blue{Di/Heike, I am not sure how I can relate this to demonstrate the power of visual inference when normal conditions are not met. Does this really test our hypothesis $H_0: \beta=0$ or some other hypothesis about the existence of cluster in the data?}}
       \label{fig:power_loess3}
\end{figure*}

@@

Model \ref{mixed} with effect as fixed covariate and subject as random effect is fitted to the data to estimate the subject specific power curves shown in figure \ref{fig:power_mixed_subject}. Subject specific power is quite different for three experiments. It appears that scatter plot serves better job compared to boxplot as a visual test statistic. But subject to subject variability is larger in scatterplot as we see in figure \ref{fig:power_mixed_subject} for experiment 2 and 3.
